---
title: "Probability Distributions"
output: pdf_document
header-includes:
  - \usepackage{lb_base_rmd}
  - \usepackage{lb_statistics}
  - \usepackage{array}
  - \usepackage{xcolor}
  - \usepackage{multicol}
  - \usepackage{multirow}
  - \newcommand{\hideFromPandoc}[1]{#1}
  - \hideFromPandoc{
      \let\Begin\begin
      \let\End\end
    }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "", fig.height=3.5)

pacman::p_load(ggplot2, ggfortify)

# How many samples to use for each distribution plot
n = 1e5

# To visualise a discrete distribution
plot_disc <- function(values, dist) {
  ggplot() + geom_col(fill = "blue", alpha=0.5, aes(x = values, y = dist(values)), data = NULL) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
}
```

\renewcommand{\arraystretch}{1.5}
\setlength\columnsep{5cm}

\newenvironment{disttab}{%
  \begin{tabular}{>{\color{cyan}}lp{6cm}}
  }
  {
  \end{tabular}
  }


### (Discrete) Uniform
\Begin{multicols}{2}

Each integer value in a specified range is assigned the same value.
\begin{disttab}
\multirow{2}{*}{Parameters} & lower bound \(a \in \Z\), \\
& upper bound \(b \in \Z\) s.t. \(a < b\) \\
Possible values & \(\set{a, a + 1, \hdots, b - 1, b}\) \\
Notation & \(U(a,b)\) \\
Probability function & \(p(x) = \frac{1}{b - a + 1}\) \\
Expectation & \(\frac{a + b}{2}\) \\
Variance & \(\frac{(b-a)(b - a + 2)}{12}\) \\
MGF & \(M(t) = \frac{1}{b - a + 1}\sum_{k=a}^b{e^{kt}}\) \\
Kernel & \(1\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,10, by=1), function(x) {return(numeric(length(x)) + 1/11)})
```

\End{multicols}

\vspace{1cm}
### Bernoulli
\Begin{multicols}{2}

Represents the success of a single experiment as a binary outcome.
\begin{disttab}
Parameters & probability of success \(0 \leq p \leq 1\) \\
Possible values & \(\set{0,1}\) \\
Notation & \(\Bern{p}\) \\
Probability function & \(p(x) = p^x(1-p)^{1-x}\) \\
Expectation & \(p\) \\
Variance & \(p(1-p)\) \\
MGF & \(M(t) = 1 + p\left(e^t - 1\right)\) \\
Kernel & \(p^x(1-p)^{-x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(c(0,1), function(x){return(0.75*as.numeric(x == 1) + 0.25*as.numeric(x == 0))})
```

\End{multicols}

\clearpage

### Binomial
\Begin{multicols}{2}
Represents the number of successes in a fixed number of independent and repeated trials of the same Bernoulli experiment. 

\begin{disttab}
\multirow{2}{*}{Parameters} & number of trials \(n \in \N\), \\
& probability of success on a single trial \(0 \leq p \leq 1\) \\
Possible values & \(\set{0,1,2,\hdots,n}\) \\
Notation & \(\Bin{n,p}\) \\
Probability function & \(p(x) = {n\choose x}p^x(1-p)^{n-x}\) \\
Expectation & \(np\) \\
Variance & \(np(1-p)\) \\
MGF & \(M(t) = \left[1 + p\left(e^t - 1\right)\right]^n\) \\
Kernel & \({n\choose x}p^x(1-p)^{-x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,10), function(x) {return(dbinom(x,p=1/3,size=10))})
```

\End{multicols}

#### Related Distributions

* If \(X \isBin{n,p}\), then 
  \[
  X = \sum_{i=1}^n{X_i},
  \]
  where the \(X_i \isBern{p}\) independently.


### Hypergeometric

\Begin{multicols}{2}
Represents the number of when drawing a fixed number of samples from a population containing a known number of successes.
\begin{disttab}
Possible values & \(\set{\max\set{0, n - (n- (N-k))},\hdots,\min\set{n,k}}\) \\
\multirow{3}{*}{Parameters} & size of population \(N \in \N\), \\
& number of successes in population \(k\in \N\), \\
& number of samples drawn \(n \in \N\) \\
Notation & \(\HypGeo{N,k,n}\) \\
Probability function & \(p(x) = \frac{{k\choose x}{N-k\choose n-x}}{{N\choose n}}\) \\
Expectation & \(\frac{nk}{N}\) \\
Variance & \(n\bfrac{k}{N}\bfrac{N-k}{N}\bfrac{N-n}{N-1}\) \\
MGF & No useful expression \\
Kernel & \({k\choose x}{N-k\choose n-x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,20), function(x) {return(dhyper(x, m = 33, n = 67, k=20))})
```

\End{multicols}

\clearpage

### Poisson

\Begin{multicols}{2}
Represents the number of events occuring in a fixed interval.
\begin{disttab}
Possible values & \(\set{0,1,2,\hdots}\) \\
\multirow{1}{*}{Parameters} & average number of events in interval \(\lambda > 0\), \\
Notation & \(\Pois{\lambda}\) \\
Probability function & \(p(x) = \frac{e^{-\lambda}\lambda^{x}}{x!}\) \\
Expectation & \(\lambda\) \\
Variance & \(\lambda\) \\
MGF & \(M(t) = e^{\lambda\left(e^t - 1\right)}\) \\
Kernel & \(\frac{\lambda^{x}}{x!}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,15), function(x) {return(dpois(x, lambda = 5))})
```

\End{multicols}


### Geometric

\Begin{multicols}{2}
Represents the number of failed Bernoulli trials preceeding the first success.
\begin{disttab}
\multirow{1}{*}{Parameters} & probability of success on a single trial \(0 \leq p \leq 1\), \\
Possible values & \(\set{0,1,2,\hdots}\) \\
Notation & \(\Geom{p}\) \\
Probability function & \(p(x) = (1-p)^xp\) \\
Expectation & \(\frac{1-p}{p}\) \\
Variance & \(\frac{1-p}{p^2}\) \\
MGF & \(M(t) = \frac{p}{1 - (1-p)e^t}\) for \(t < -\ln(1-p)\) \\
Kernel & \((1-p)^x\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,20), function(x) {return(dgeom(x, prob = 0.25))})
```

\End{multicols}

\clearpage

### Geometric (alternative)

\Begin{multicols}{2}
Represents the position of the first success in a sequence of Bernoulli trials.
\begin{disttab}
\multirow{1}{*}{Parameters} & probability of success on a single trial \(0 \leq p \leq 1\), \\
Possible values & \(\set{1,2,\hdots}\) \\
Notation & \(\Geom{p}\) \\
Probability function & \(p(x) = (1-p)^{x-1}p\) \\
Expectation & \(\frac1p\) \\
Variance & \(\frac{1-p}{p^2}\) \\
MGF & \(M(t) = \frac{pe^t}{1 - (1-p)e^t}\) for \(t < -\ln(1-p)\) \\
Kernel & \((1-p)^{x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(1,20), function(x) {return(dgeom(x-1, prob = 0.25))})
```

\End{multicols}



### Negative Binomial

\Begin{multicols}{2}
Represents the number of Bernoulli trials preceeding the \(r\)th success.
\begin{disttab}
\multirow{2}{*}{Parameters} & number of desired successes \(r \in \N\), \\ 
& probability of success on a single trial \(0 \leq p \leq 1\), \\
Possible values & \(\set{r,r+1,\hdots}\) \\
Notation & \(\NegBin{r,p}\) \\
Probability function & \(p(x) = {x-1 \choose r-1}(1-p)^{x-r}p^r\) \\
Expectation & \(\frac{r}{p}\) \\
Variance & \(\frac{r(1-p)}{p^2}\) \\
MGF & \(M(t) = \left[\frac{pe^t}{1 - (1-p)e^t}\right]^r\) for \(t < -\ln{p}\) \\
Kernel & \({x-1 \choose r-1}(1-p)^{x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(5,40), function(x) {return(dnbinom(x, size = 5, prob = 0.25))})
```

\End{multicols}

#### Related Distributions
* If \(X \isNegBin{r,p}\), then 
  \[
  X = \sum_{i=1}^n{X_i},
  \]
  where the \(X_i \isGeom{p}\) independently.


\clearpage

### Negatve Binomial (Alternative)

\Begin{multicols}{2}
Represents the number of failures preceeding the \(r\)th success in a sequence of Bernoulli trials.
\begin{disttab}
\multirow{2}{*}{Parameters} & number of desired successes \(r \in \N\), \\ 
& probability of success on a single trial \(0 \leq p \leq 1\), \\
Possible values & \(\set{0,1,2,\hdots}\) \\
Notation & \(\NegBin{r,p}\) \\
Probability function & \(p(x) = {x+r-1 \choose x}(1-p)^{x}p^r\) \\
Expectation & \(\frac{r(1-p)}{p}\) \\
Variance & \(\frac{r(1-p)}{p^2}\) \\
MGF & \(M(t) = \left[\frac{p}{1 - (1-p)e^t}\right]^r\) for \(t < -\ln{p}\) \\
Kernel & \({x+r-1 \choose x}(1-p)^{x}\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(5,40), function(x) {return(dnbinom(x, size = 5, prob = 0.25))})
```

\End{multicols}


## Continuous

### (Continuous) Uniform
\Begin{multicols}{2}

Each value in a specified interval has the same probability density.
\begin{disttab}
\multirow{2}{*}{Parameters} & lower bound \(a \in \R\), \\
& upper bound \(b \in \R\) s.t. \(a < b\) \\
Possible values & \((a,b)\) \\
Notation & \(\Uniform{a,b}\) \\
Probability density function & \(f(x) = \frac{1}{b - a}\) \\
Cumulative distribution function & \(F(x) = \begin{cases}0, & \text{if } x \leq a, \\ \frac{x-a}{b-a}, & \text{if } a < x < b, \\ 1,& \text{if } b \leq x.\end{cases}\) \\
Expectation & \(\frac{a + b}{2}\) \\
Variance & \(\frac{(b-a)^2}{12}\) \\
MGF & \(M(t) = \begin{cases}1, & \text{if } t = 0, \\ \frac{e^{tb}-e^{ta}}{t(b-a)}, & \text{if } t \neq 0.\end{cases}\) \\
Kernel & \(1\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dunif, seq(0,10, by=1), min = 0, max = 10, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

\clearpage

### Exponential
\Begin{multicols}{2}

Represents the waiting time until an event occurs.
\begin{disttab}
\multirow{1}{*}{Parameters} & rate of events \(\lambda > 0\) \\
Possible values & \((0,\infty)\) \\
Notation & \(\Exp{\lambda}\) \\
Probability density function & \(f(x) = \lambda e^{-\lambda x}\) \\
Cumulative distribution function & \(F(x) = \begin{cases}0, & \text{if } x \leq 0, \\ 1 - e^{-\lambda x}, & \text{if } x > 0.\end{cases}\) \\
Expectation & \(\frac{1}{\lambda}\) \\
Variance & \(\frac{1}{\lambda^2}\) \\
MGF & \(M(t) = \frac{\lambda}{\lambda - t}\) for \(t < \lambda\) \\
Kernel & \(e^{-\lambda x}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dexp, seq(0,5, by=0.1), rate=1, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}



### Exponential (Alternative)
\Begin{multicols}{2}

Characterised by average time until the first event \(\beta = 1/\lambda\).
\begin{disttab}
\multirow{1}{*}{Parameters} & average time until event \(\beta > 0\) \\
Possible values & \((0,\infty)\) \\
Notation & \(\Exp{\beta}\) \\
Probability density function & \(f(x) = \frac{1}{\beta}e^{-\frac{x}{\beta}}\) \\
Cumulative distribution function & \(F(x) = \begin{cases}0, & \text{if } x \leq 0, \\ 1 - e^{-\frac{x}{\beta}}, & \text{if } x > 0.\end{cases}\) \\
Expectation & \(\beta\) \\
Variance & \(\beta^2\) \\
MGF & \(M(t) = \frac{1}{1 - \beta t}\) for \(t < \frac1\beta\) \\
Kernel & \(e^{-\frac{x}{\beta}}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dexp, seq(0,5, by=0.1), rate=1, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}


### Gamma
\Begin{multicols}{2}

Generalisation of the exponential distribution. For integer shape parameter, represents the waiting time until the \(r\)th event.
\begin{disttab}
\multirow{2}{*}{Parameters} & shape \(r>0\) \\
& scale \(\lambda > 0\) \\
Possible values & \((0,\infty)\) \\
Notation & \(\GamDist{a,b}\) \\
Probability density function & \(f(x) = \frac{\lambda^r}{\Gamma(r)}x^{r-1}e^{-\lambda x}\) \\
Cumulative distribution function & no simple expression \\
Expectation & \(\frac{r}{\lambda}\) \\
Variance & \(\frac{r}{\lambda^2}\) \\
MGF & \(M(t) = \bfrac{\lambda}{\lambda - t}^r\) for \(t < \lambda\) \\
Kernel & \(x^{r-1}e^{-\lambda x}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dgamma, seq(0,15, by=0.1), shape=5, rate=1, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

#### Properties

* If \(X \isGamma{r,\lambda}\), then for any \(k > 0\), \(kX \isGamma{r,\frac{\lambda}{k}}\)

#### Related Distributions

* If \(r \in \N\) and \(X \isGamma{r,\lambda}\), then
\[
X = \sum_{i=1}^n{X_i},
\]
where \(X_i \isExp{\lambda}\) independently.

* If \(X \isChiSq{r}\), then \(X \isGamma{\frac{r}{2}, \frac12}\).



### Beta
\Begin{multicols}{2}

Represents a random variable which is positive and bounded. Useful for modelling proportions and percentages.
\begin{disttab}
\multirow{2}{*}{Parameters} & shape \(\alpha>0\) \\
& shape \(\beta > 0\) \\
Possible values & \((0,1)\) \\
Notation & \(\BetaDist{a,b}\) \\
Probability density function & \(f(x) = \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}x^{\alpha-1}(1-x)^{\beta-1}\) \\
Cumulative distribution function & no simple expression \\
Expectation & \(\frac{\alpha}{\alpha + \beta}\) \\
Variance & \(\frac{\alpha\beta}{(\alpha+\beta+1)(\alpha+\beta)^2}\) \\
MGF & exists but no simple expression \\
Kernel & \(x^{\alpha-1}(1-x)^{\beta-1}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dbeta, seq(0,1, by=0.01), shape1=3, shape2=6, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

#### Related Distributions

* If \(X \isGamma{\alpha,\lambda}\) and \(Y \isGamma{\beta,\lambda}\) independently, then \(\frac{X}{X + Y} \isBeta{\alpha,\beta}\).

* The Beta distribution can be generalised to any bounded interval \([a,b]\), with properties

  - 
  
  
  

### Normal
\Begin{multicols}{2}

The classic bell-shaped curve.
\begin{disttab}
\multirow{2}{*}{Parameters} & mean \(\mu \in \R\) \\
& variance \(\sigma^2 > 0\) \\
Possible values & \(\R\) \\
Notation & \(\Gauss{\mu,\sigma^2}\) \\
Probability density function & \(f(x) = \frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{(x-\mu)^2}{2\sigma^2}}\) \\
Cumulative distribution function & no simple expression \\
Expectation & \(\mu\) \\
Variance & \(\sigma^2\) \\
MGF & \(e^{\mu t + \frac{\sigma^2t^2}{2}}\) \\
Kernel & \(e^{-\frac{(x-\mu)^2}{2\sigma^2}}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dnorm, seq(-3,3, by=0.1), colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}


### \(\chi^2\)
\Begin{multicols}{2}

The sum of the squares of \(n\) independent standard normal random variables. 
\begin{disttab}
\multirow{1}{*}{Parameters} & degrees of freedom \(n \in \N\) \\
Possible values & \((0,\infty)\) if \(n = 1\); \([0,\infty)\) otherwise \\
Notation & \(\ChiSq{n}\) \\
Probability density function & \(f(x) = \frac{1}{2^{n/2}\Gamma\bfrac{n}{2}}x^{\frac{n}{2}-1}e^{-\frac{x}{2}}\) \\
Cumulative distribution function & \(F(x) = \begin{cases}0, & \text{if } x \leq 0, \\ \frac{1}{\Gamma\bfrac{n}{2}}\int_{0}^{x/2}{t^{\frac{n}{2} - 1}e^{-\frac{t}{2}}\dif t}. & \text{if } x > 0\end{cases}\) \\
Expectation & \(n\) \\
Variance & \(2n\) \\
MGF & \((1-2t)^{-\frac{n}{2}}\) \\
Kernel & \(x^{\frac{n}{2}-1}e^{-\frac{x}{2}}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dchisq, seq(0,25, by=0.1), df=10, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

#### Related Distributions

* If \(X \isChiSq{n}\), then 
\[
X = \sum_{i=1}^{n}{Z_i^2},
\]
where \(Z_i \isStdGauss\) independently.

* If \(X \isChiSq{r}\), then \(X \isGamma{\frac{r}{2}, \frac12}\).

* If \(Y \isF{\nu_1, \nu_2}\), then \(\lim_{\nu_2\to\infty}\nu_1Y \isChiSq{\nu_1}\).


### Cauchy
The ratio of two independent, identically distributed and central normal random variables.

\Begin{multicols}{2}

The sum of the squares of \(n\) independent standard normal random variables. 
\begin{disttab}
\multirow{2}{*}{Parameters} & location \(\mu \in \R\) \\
& scale \(\sigma > 0\) \\
Possible values & \(\R\) \\
Notation & \(\Cauchy{\mu,\sigma}\) \\
Probability density function & \(f(x) = \frac{1}{\pi\sigma\left(1 + \frac{2(x - \mu)}{\sigma}\right)}\) \\
Cumulative distribution function & \(F(x) = \begin{cases}0, & \text{if } x \leq 0, \\ \frac12 + \frac1\pi\arctan\bfrac{x - \mu}{\sigma}, & \text{if } x > 0\end{cases}\) \\
Expectation & does not exist \\
Variance & does not exist \\
MGF & does not exist \\
Kernel & \(\sigma + 2(x-\mu)\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dcauchy, seq(-5,5, by=0.1), colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

#### Related Distributions

* If \(Z_1, Z_2 \isStdGauss\) independently, then \(Z_1/Z_2 \isCauchy{0,1}\).

* If \(X \isCauchy{\mu,\sigma}\) then \(X \isT{n}{\mu,\sigma}\).


### \(t\)
\Begin{multicols}{2}

The ratio of a standard normal random variable and \(\chi^2\) random variable.
\begin{disttab}
\multirow{3}{*}{Parameters} & location \(\mu \in \R\) \\
& scale \(\sigma > 0\) \\
& degrees of freedom \(\nu \in \N\) \\
Possible values & \((0,\infty)\) if \(n = 1\); \([0,\infty)\) otherwise \\
Notation & \(\TDist{n}{\mu,\sigma^2}\) \\
Probability density function & \(f(x) = \frac{\Gamma\bfrac{\nu+1}{2}}{\Gamma\bfrac{\nu}{2}\sigma\sqrt{\pi\nu}}\left[1 + \frac1\nu\bfrac{x-\mu}{\sigma}^2\right]^{-\frac{\nu+1}{2}}\) \\
Cumulative distribution function & no simple expression \\
Expectation & \(\mu\) for \(\nu\geq 2\) \\
Variance & \(\frac{\sigma^2\nu}{\nu-2}\) for \(\nu\leq 3\) \\
MGF & does not exist \\
Kernel & \(\left[1 + \frac1\nu\bfrac{x-\mu}{\sigma}^2\right]^{-\frac{\nu+1}{2}}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(dt, seq(-5,5, by=0.1), df=10, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}

#### Related Distributions

* If \(Z\isStdGauss\) and \(W\isChiSq{\nu}\) independently, then 
\[
\mu + \sigma\frac{Z}{\sqrt{\frac{W}{\nu}}} \isT{\nu}{\mu,\sigma^2}.
\]


### \(F\)
\Begin{multicols}{2}

The ratio of the sample variance of two normal random variables.
\begin{disttab}
\multirow{2}{*}{Parameters} & degrees of freedom \(\nu_1 \in \N\) \\
& degrees of freedom \(\nu_2 \in \N\) \\
Possible values & \((0,\infty)\) \\
Notation & \(\FDist{n}{\mu,\sigma^2}\) \\
Probability density function & \(f(x) = \frac{\Gamma\bfrac{\nu_1\nu_2}{2}\nu_1^{\nu_1/2}\nu_2^{\nu_2}x^{\nu_1/2 - 1}}{\Gamma\bfrac{\nu_1}{2}\Gamma{\nu_2}{2}\left(\nu_1x + \nu_2\right)^{\frac{\nu_1 + \nu_2}{2}}}\) \\
Cumulative distribution function & no simple expression \\
Expectation & \(\frac{\nu_2}{\nu_2 - 2}\) for \(\nu\geq 2\) \\
Variance & \(\frac{2\nu_2^2(\nu_1 + \nu_2 - 2)}{\nu_1(\nu_2 - 2)^2(\nu_2 - 4)}\) for \(\nu\geq 4\) \\
MGF & does not exist \\
Kernel & \(\frac{x^{\nu_1/2 - 1}}{\left(\nu_1x + \nu_2\right)^{\frac{\nu_1 + \nu_2}{2}}}\)
\end{disttab}

\columnbreak

```{r}
ggdistribution(df, seq(0,5, by=0.01), df1=10, df2 = 8, colour = "blue", fill = "blue", alpha=0.5) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\End{multicols}


#### Related Distributions

* If \(U \isChiSq{\nu_1}\) and \(V\isChiSq{\nu_2}\) independently, then
\[
\frac{U/\nu_1}{V/\nu_2} \isF{\nu_1, \nu_2}.
\]

* If \(X_1 \isGamma{\alpha_1,\beta_1}\) and \(X_2\isGamma{\alpha_2,\beta_2}\) independently, then
\[
\frac{\alpha_1\beta_1X_1}{\alpha_2\beta_2X_2} \isF{2\alpha_1, 2\alpha_2}.
\]

* If \(X \isBeta{\alpha/2, \beta/2}\) then 
\[
\frac{\beta X}{\alpha(1 - X)} \isF{\alpha,\beta}.
\]

* If \(X \isStdT{\nu}\), then \(X^2 \isF{1,\nu}\).






# Multivariate


## Discrete


## Continuous