---
title: "Probability Distributions"
output: pdf_document
header-includes:
  - \usepackage{lb_base_rmd}
  - \usepackage{lb_statistics}
  - \usepackage{array}
  - \usepackage{xcolor}
  - \usepackage{multicol}
  - \usepackage{multirow}
  - \newcommand{\hideFromPandoc}[1]{#1}
  - \hideFromPandoc{
      \let\Begin\begin
      \let\End\end
    }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "", fig.height=3.5)

pacman::p_load(ggplot2)

# How many samples to use for each distribution plot
n = 1e5

# To visualise a discrete distribution
plot_disc <- function(values, dist) {
  ggplot() + geom_col(fill = "blue", alpha=0.5, aes(x = values, y = dist(values)), data = NULL) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
          axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
}
```

\renewcommand{\arraystretch}{1.5}
\setlength\columnsep{5cm}

\newenvironment{disttab}{%
  \begin{tabular}{>{\color{cyan}}lp{5cm}}
  }
  {
  \end{tabular}
  }


### (Discrete) Uniform
\Begin{multicols}{2}

Each integer value in a specified range is assigned the same value.
\begin{disttab}
\multirow{2}{*}{Parameters} & lower bound \(a \in \Z\), \\
& upper bound \(b \in \Z\) s.t. \(a < b\) \\
Possible values & \(\set{a, a + 1, \hdots, b - 1, b}\) \\
Notation & \(U(a,b)\) \\
Probability function & \(p(x) = \frac{1}{b - a + 1}\) \\
Expectation & \(\frac{a + b}{2}\) \\
Variance & \(\frac{(b-a)(b - a + 2)}{12}\) \\
MGF & \(M(t) = \frac{1}{b - a + 1}\sum_{k=a}^b{e^{kt}}\) \\
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,10, by=1), function(x) {return(numeric(length(x)) + 1/11)})
```

\End{multicols}

\vspace{1cm}
### Bernoulli
\Begin{multicols}{2}

Represents the success of a single experiment as a binary outcome.
\begin{disttab}
Parameters & probability of success \(0 \leq p \leq 1\) \\
Possible values & \(\set{0,1}\) \\
Notation & \(\Bern{p}\) \\
Probability function & \(p(x) = p^x(1-p)^{1-x}\) \\
Expectation & \(p\) \\
Variance & \(p(1-p)\) \\
MGF & \(M(t) = 1 + p\left(e^t - 1\right)\)
\end{disttab}

\columnbreak

```{r}
plot_disc(c(0,1), function(x){return(0.75*as.numeric(x == 1) + 0.25*as.numeric(x == 0))})
```

\End{multicols}

\clearpage

### Binomial
\Begin{multicols}{2}
Represents the number of successes in a fixed number of independent and repeated trials of the same Bernoulli experiment. 

\begin{disttab}
\multirow{2}{*}{Parameters} & number of trials \(n \in \N\), \\
& probability of success on a single trial \(0 \leq p \leq 1\) \\
Possible values & \(\set{0,1,2,\hdots,n}\) \\
Notation & \(\Bin{n,p}\) \\
Probability function & \(p(x) = {n\choose x}p^x(1-p)^{n-x}\) \\
Expectation & \(np\) \\
Variance & \(np(1-p)\) \\
MGF & \(M(t) = \left[1 + p\left(e^t - 1\right)\right]^n\)
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,10), function(x) {return(dbinom(x,p=1/3,size=10))})
```

\End{multicols}

#### Related Distributions

* If \(X \isBin{n,p}\), then 
  \[
  X = \sum_{i=1}^n{X_i},
  \]
  where the \(X_i\) are independent and identically distributed random variables with probability \(p\) of success.


### Hypergeometric

\Begin{multicols}{2}
Represents the number of when drawing a fixed number of samples from a population containing a known number of successes.
\begin{disttab}
Possible values & \(\set{\max\set{0, n - (n- (N-k))},\hdots,\min\set{n,k}}\) \\
\multirow{3}{*}{Parameters} & size of population \(N \in \N\), \\
& number of successes in population \(k\in \N\), \\
& number of samples drawn \(n \in \N\)
Notation & \(\HypGeo{N,k,n}\) \\
Probability function & \(p(x) = \frac{{k\choose x}{N-k\choose n-x}}{{N\choose n}}\) \\
Expectation & \(\frac{nk}{N}\) \\
Variance & \(n\bfrac{k}{N}\bfrac{N-k}{N}\bfrac{N-n}{N-1}\) \\
MGF & No useful expression
\end{disttab}

\columnbreak

```{r}
plot_disc(seq(0,20), function(x) {return(dhyper(x, m = 33, n = 67, k=20))})
```

\End{multicols}




## Continuous


# Multivariate


## Discrete


## Continuous